---
title: "Homework 3"
author: "Kelvin Delali Tamakloe"
date: 10/20/2021
output: github_document
---
```{r setup}
library(tidyverse)
library(dplyr)
library(p8105.datasets)
library(janitor)
library(readr)
```

# Question 1

## Instacart Dataset Exploration

Below, we will explore the instacart dataset

```{r}
data("instacart")
```

Brief Description:

Instacart is an American grocery delivery and pick up service company. 

The `instacart` dataset has data from the year 2017. There are `r nrow(instacart)` observations and `r ncol(instacart)` variables. 

In this dataset, the the total number of data points is `r nrow(instacart)*ncol(instacart)`. There are some key variables in this dataset including `r colnames(instacart)`.

## Total number of aisles and aisles from which the most items are ordered

The code below will tell us how many aisles there are, and from which aisles the most items are ordered.

```{r, message = FALSE, warning = FALSE}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```
There are a total 134 `aisles`. The aisles from which the most items are ordered are `fresh vegetables`, `fresh fruits`, and `packaged vegetables fruits` in that order.

## Plot of number of items ordered per aisle (> 10000 items ordered)

The code below generates a plot of the number of items ordered in each aisle but limited to aisles with more than 10000 orders.

```{r aisles_plot}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  ggplot(aes(forcats::fct_reorder(aisle, (n)), n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of items ordered per aisle",
    x = "Aisle name",
    y = "Number of items ordered",
    caption = "Data from instacart limited to aisles with more than 10,000 items ordered"
  )
```

## Table showing three most popular items in selected aisles 
The code below  produces a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”,with information on the number of items ordered.

```{r aisles_table}
instacart %>%
  filter(aisle == "baking ingredients" | 
           aisle == "dog food care" | 
           aisle == "packaged vegetables fruits") %>%
 count(aisle, product_name) %>%
  arrange(desc(n)) %>%
  group_by(aisle) %>%
  slice(1:3) %>%
  arrange(desc(n)) %>%
  group_by(aisle) %>%
  knitr::kable(caption = "Three Most Ordered Items Per Aisle")
```

From the table, `organic baby spinach`, `organic raspberries`, and `organic blueberries` 
are the three most ordered items from the `packaged vegetables fruits` aisle. 

Also, `light brown sugar`, `pure baking soda`, and `cane sugar` are the three most ordered items from the `baking ingredients` aisle. 

The three most ordered items from the `dog food care` aisle are `snack sticks chicken & rice recipe dog treats`, `organix chicken & brown rice recipe`, and `small dog biscuits`.


## Table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r apples_ice_cream_table}
 instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
  ) %>%
  rename( "Sunday" = "0","Monday" = "1", "Tuesday" = "2", "Wednesday" = "3", "Thursday" = "4", "Friday" = "5", "Saturday" = "6") %>% 
  knitr::kable(caption = "Mean Hour of Day Ordered")
```
Generally, `Pink Lady Apples` are ordered earlier in the day than `Coffee Ice Cream`.

# Question 2

The code below will clean the `brfss_smart2010` data: formatting it to use appropriate variable names, 
focusing on the `Overall Health` topic, and only including factored responses ordered from`Poor` to `Excellent`.

```{r clean_brfss}
data("brfss_smart2010")
brfss_cleaned_version = brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic %in% c("Overall Health")) %>%
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>%
  arrange(response) %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  separate(locationdesc, into = c('state', 'location'), sep = ' - ')
```

## 2002: States observed at 7 or more locations

The code below will provide information on which states were observed at 7 or more locations in 2002.

```{r states2002}
states_2002 = brfss_cleaned_version %>% 
  filter(year == "2002") %>% 
  group_by(state) %>% 
  distinct(location) %>% 
  count(state) %>% 
  filter(n >= 7) %>% 
  select(state)
```

## 2010: States observed at 7 or more locations

The code below will provide information on which states were observed at 7 or more locations in 2002.

```{r states2010}
states_2010 = brfss_cleaned_version %>% 
  filter(year == "2010") %>% 
  group_by(state) %>% 
  distinct(location) %>% 
  count(state) %>% 
  filter(n >= 7) %>% 
  select(state)
```

In 2002, the states that were observed at 7 or more locations were `r states_2002$state`. 
In 2010, the states that were observed at 7 or more locations were `r states_2010$state`.