---
title: "Homework 3"
author: "Kelvin Delali Tamakloe"
date: 10/20/2021
output: github_document
---
```{r setup}
library(tidyverse)
library(dplyr)
library(p8105.datasets)
library(janitor)
library(readr)
```

# Question 1 

Below, we will explore the instacart dataset

```{r}
data("instacart")
```

Brief Description:
Instacart is an American grocery delivery and pick up service company. The `instacart` dataset has data from the year 2017. There are `r nrow(instacart)` observations and `r ncol(instacart)` variables. In this dataset, the the total number of data points `r nrow(instacart)*ncol(instacart)`. There are some key variables in this dataset including `r colnames(instacart)`.

## Total number of aisles and aisles from which the most items are ordered

The code below will tell us how many aisles there are, and from which aisles the most items are ordered.

```{r, message = FALSE, warning = FALSE}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```
There are a total 134 `aisles`. The aisles from which the most items are ordered are `fresh vegetables`, `fresh fruits`, and followed by `packaged vegetables fruits`. 